GPU-friendly Stroke Expansion

### 2024-06-13 paper.tex
Tweaks to flatten algorithm
Small changes to the pseudocode, and fixing up the text to be more consistent with it.

Correctness is another challenging aspect of stroke rendering. We propose that the correctness of stroke outlines be divided into \emph{weak correctness} and \emph{strong correctness.} We define strong correctness as the computation of the outline of a line swept along the segment, maintaining normal orientation, combined with stroke caps and joins. Weak correctness, by contrast, only requires the parallel curves of path segments, combined with caps and the outer contours of joins. Both \citet{Nehab2020} and \citet{Kilgard2020} provide useful definitions of strongly correct stroke rendering, and Nehab in particular describes how to achieve it in stroke expansion. Briefly, when the path curvature exceeds the reciprocal of the stroke half-width, then addition \emph{evolute} segments need to be added, in addition to the parallel curves required by weak correctness, and also the inner contours of joins are required in general. The technique described in this paper is intended for strongly correct stroke rendering; it produces outlines that render the same shape as \citet{Nehab2020}.

@@

\subsection{Optimal flattening}

The optimal flattening of a curve contains the minimum number of subdivisions required to satisfy a maximum error constraint, in which the error is the minimum possible with that number of subdivisions.

For a convex curve, the optimal flattening is tractable to compute, though not necessarily very fast. When a curve is monotonic, error is also monotonic with respect to both subdivision points. Thus, any standard root finding technique can find a subdivision with a given error (bisection is the most popular, but the ITP method[\citenum{Oliveira2020}] is better). A similar outer loop with a root finding technique can find the error which minimizes the global error for the entire flattening, which, due to monotonicity, is that for which the error of the final segment equals that of the previous segments. Subdividing at the inflection point adds at most one additional point; it is possible to optimize further, but we do not consider that (though the near-optimal flattening we will later present is not subject to this constraint).

@@

The entire recursion stack for adaptive subdivision can be represented in two words: $dt$ and $\mbox{scaled}\_t0$. The range is $dt \cdot \mbox{scaled}\_t0 .. dt \cdot (\mbox{scaled}\_t0 + 1)$. Initial values of $1$ and $0$ represent the range $(0, 1)$. Pushing the stack, subdividing the range in half, is represented by halving $dt$ and doubling $\mbox{scaled}\_t0$, which leaves the start of the range invariant but its size halved. After accepting an approximation, the next range is determined by incrementing $\mbox{scaled}\_t0$, then repeatedly popping the stack by doubling $dt$ and halving $\mbox{scaled}\_t0$, as long as the latter is odd. This can in fact be achieved without iteration by using the ``count trailing zeros'' intrinsic.

@@

We applied the techniques outlined in Section 7 to implement a data-parallel algorithm that can convert stroked 2D Bézier paths into flat geometry suitable for GPU rendering. We focused primarily on the global stroke-to-fill conversion problem in order to generate polygonal outlines for rasterization. Since the Euler spiral approximation can fit any source cubic Bézier and its parallel offset curves, our implementation can be used to render both filled and stroked primitives within a satisfactory error tolerance.

@@

\begin{center}
\begin{figure}
    \includegraphics[scale=0.18]{stroke_threads}
    \caption{Visualization of contour outlines and the IDs of GPU threads that flatten them. The subpath contains two adjacent cubic Bézier curves rendered as both a stroke and a fill. In the stroked version, thread $n$ flattens the offset curves of the first curve (blue) and the join segment with the miter style (dark blue). Thread $n + 1$ flattens the offset curves of the second curve (orange) and the end cap with the round style (dark orange). Thread $n + 2$ handles the \emph{subpath end segment} and flattens the start cap (red). When the same path is rendered as a fill, threads $n$ and $n + 1$ each flatten the source curves, while thread $n + 2$ outputs the \emph{close} segment that joins the end of the subpath to its starting point. The black arrows represent the winding direction of the output segments.}
    \Description{TODO}
    \label{fig:stroke_threads}
\end{figure}
\end{center}

\begin{algorithm}
  $tid \gets$ global invocation ID\;
  $(style, segment) \gets readScene(tid)$\;
  \eIf{$style$ is $fill$}{
      \tcp{Output source curve}
      flatten\_cubic(segment, $0$)\;
  }{
      \tcp{Handle stroked segment}
      \eIf{$segment$ is \emph{stroke cap marker}}{
          \eIf{path is open}{output start cap\;}{output nothing\;}
      }{
          \tcp{Output both offset curves}
          flatten\_cubic(segment,  style.offset)\;
          flatten\_cubic(segment, -style.offset)\;

@@

\begin{algorithm}
\SetKwProg{Fn}{fn}{}{end}
\SetKwBlock{Loop}{loop}{end}
\Fn{flatten\_cubic(cubic: Cubic, offset: f32)}{
    TOL $\gets$ Euler spiral fit tolerance\;
    \tcp{'t0\_u' and 'dt' track the recursion stack}
    t0\_u $\gets$ 0u\;
    dt $\gets$ 1.0\;
    last\_p $\gets$ cubic.p0\;
    last\_q $\gets$ cubic.p1 - cubic.p0\;
    \Loop{
        t0 $\gets$ f32(t0\_u) * dt\;
        \If{dt == 1.0}{break\;}
        t1 $\gets$ t0 + dt\;
        (next\_p, next\_q) $\gets$ eval\_cubic\_and\_deriv(cubic, t1)\;
        \tcp{Estimate error of geometric}
        \tcp{Hermite interpolation to Euler spiral}
        \nosemic(th0, th1, chord\_len, err) $\gets$ fit\_cubic\_to\_es(\;
        \pushline\dosemic last\_p, next\_p, last\_q, next\_q, dt)\;
        \popline \eIf{err <= TOL}{
            es $\gets$ euler\_segment(th0, th1, last\_p, next\_p)\;
            \tcp{Find location of any cusp in outline.}
            \tcp{If $0 \leq t < 1$, draw evolute at 't'.}
            cusp0 $\gets$ es\_curvature(es, 0) * offset + chord\_len\;
            cusp1 $\gets$ es\_curvature(es, 1) * offset + chord\_len\;
            t $\gets$ cusp0 * cusp1 >= 0 ? 1 : cusp0 / (cusp0 - cusp1)\;
            \If{cusp0 >= 0}{
                evolute\_finalize()\;
                es\_seg\_flatten\_offset(es, offset, vec2(0, t))\;
            }
            \If{cusp0 < 0 OR t < 1}{
                evolute\_start(es, offset)\;
                range $\gets$ cusp0 >= 0 ? vec2(t, 1) : vec2(0, t)\;
                es\_seg\_flatten\_evolute(es, range)\;
                es\_seg\_flatten\_offset\_reverse(es, offset, range)\;
            }
            \If{cusp0 < 0 AND t < 1}{
                evolute\_finalize()\;
                es\_seg\_flatten\_offset(es, offset, vec2(t, 1))\;
            }
            \tcp{Pop the stack}
            t0\_u $\gets$ t0\_u + 1u\;
            shift $\gets$ countTrailingZeros(t0\_u)\;
            t0\_u $\gets$ t0\_u >> shift;
            dt $\gets$ dt * f32(1u << shift)\;
            last\_p $\gets$ next\_p\;
            last\_q $\gets$ next\_q\;
        }{
            \tcp{Error is above the tolerance.}
            \tcp{Push the stack; Continue subdivision}
            t0\_u $\gets$ t0\_u * 2u\;
            dt $\gets$ dt * 0.5\;
        }
    }
}
\caption{Outline of the iterative cubic Bézier flattening routine invoked in our compute shader. A working WGSL implementation is provided in the supplemental materials.}
\label{alg:flatten-cubic}
\end{algorithm}

>>>>

Correctness is another challenging aspect of stroke rendering. We propose that the correctness of stroke outlines be divided into \emph{weak correctness} and \emph{strong correctness.} We define strong correctness as the computation of the outline of a line swept along the segment, maintaining normal orientation, combined with stroke caps and joins. Weak correctness, by contrast, only requires the parallel curves (also known as ``offset curves'') of path segments, combined with caps and the outer contours of joins. Both \citet{Nehab2020} and \citet{Kilgard2020} provide useful definitions of strongly correct stroke rendering, and Nehab in particular describes how to achieve it in stroke expansion. Briefly, when the path curvature exceeds the reciprocal of the stroke half-width, then addition \emph{evolute} segments need to be added, in addition to the parallel curves required by weak correctness, and also the inner contours of joins are required in general. The technique described in this paper is intended for strongly correct stroke rendering; it produces outlines that render the same shape as \citet{Nehab2020}.

@@

%\subsection{Optimal flattening}
%
%The optimal flattening of a curve contains the minimum number of subdivisions required to satisfy a maximum error constraint, in which the error is the minimum possible with that number of subdivisions.
%
%For a convex curve, the optimal flattening is tractable to compute, though not necessarily very fast. When a curve is monotonic, error is also monotonic with respect to both subdivision points. Thus, any standard root finding technique can find a subdivision with a given error (bisection is the most popular, but the ITP method[\citenum{Oliveira2020}] is better). A similar outer loop with a root finding technique can find the error which minimizes the global error for the entire flattening, which, due to monotonicity, is that for which the error of the final segment equals that of the previous segments. Subdividing at the inflection point adds at most one additional point; it is possible to optimize further, but we do not consider that (though the near-optimal flattening we will later present is not subject to this constraint).

@@

The entire recursion stack for adaptive subdivision can be represented in two words: $dt$ and $t0\_u$ (the latter being an unsigned integer). The range is $dt \cdot t0\_u .. dt \cdot (t0\_u + 1)$. Initial values of $1$ and $0$ represent the range $(0, 1)$. Pushing the stack, subdividing the range in half, is represented by halving $dt$ and doubling $t0\_u$, which leaves the start of the range invariant but its size halved. After accepting an approximation, the next range is determined by incrementing $t0\_u$, then repeatedly popping the stack by doubling $dt$ and halving $t0\_u$, as long as the latter is odd. This can in fact be achieved without iteration by using the ``count trailing zeros'' intrinsic.

@@

We applied the techniques outlined in Section 7 to implement a data-parallel algorithm that can convert stroked 2D Bézier paths into flat geometry suitable for GPU rendering. We focused primarily on the global stroke-to-fill conversion problem in order to generate polygonal outlines for rasterization. Since the Euler spiral approximation can fit any source cubic Bézier and its parallel curves, our implementation can be used to render both filled and stroked primitives within a satisfactory error tolerance.

@@

\begin{center}
\begin{figure}
    \includegraphics[scale=0.18]{stroke_threads}
    \caption{Visualization of contour outlines and the IDs of GPU threads that flatten them. The subpath contains two adjacent cubic Bézier curves rendered as both a stroke and a fill. In the stroked version, thread $n$ flattens the parallel curves of the first curve (blue) and the join segment with the miter style (dark blue). Thread $n + 1$ flattens the parallel curves of the second curve (orange) and the end cap with the round style (dark orange). Thread $n + 2$ handles the \emph{subpath end segment} and flattens the start cap (red). When the same path is rendered as a fill, threads $n$ and $n + 1$ each flatten the source curves, while thread $n + 2$ outputs the \emph{close} segment that joins the end of the subpath to its starting point. The black arrows represent the winding direction of the output segments.}
    \Description{TODO}
    \label{fig:stroke_threads}
\end{figure}
\end{center}

\begin{algorithm}
  $tid \gets$ global invocation ID\;
  $(style, segment) \gets readScene(tid)$\;
  \eIf{$style$ is $fill$}{
      \tcp{Output source curve}
      flatten\_cubic(segment, $0$)\;
  }{
      \tcp{Handle stroked segment}
      \eIf{$segment$ is \emph{stroke cap marker}}{
          \eIf{path is open}{output start cap\;}{output nothing\;}
      }{
          \tcp{Output both parallel curves}
          flatten\_cubic(segment,  style.offset)\;
          flatten\_cubic(segment, -style.offset)\;

@@

\begin{algorithm}
\SetKwProg{Fn}{fn}{}{end}
\SetKwBlock{Loop}{loop}{end}
\Fn{flatten\_cubic(cubic: Cubic, offset: f32)}{
    TOL $\gets$ Euler spiral fit tolerance\;
    \tcp{'t0\_u' and 'dt' track the recursion stack}
    t0\_u $\gets$ 0\;
    dt $\gets$ 1.0\;
    last\_p $\gets$ cubic.p0\;
    last\_q $\gets$ cubic.p1 - cubic.p0\;
    \Loop{
        t0 $\gets$ f32(t0\_u) * dt\;
        \If{t0 == 1.0}{break\;}
        t1 $\gets$ t0 + dt\;
        (next\_p, next\_q) $\gets$ eval\_cubic\_and\_deriv(cubic, t1)\;
        \tcp{Estimate error of geometric}
        \tcp{Hermite interpolation to Euler spiral}
        \nosemic(th0, th1, chord\_len, err) $\gets$ fit\_cubic\_to\_es(\;
        \pushline\dosemic last\_p, next\_p, last\_q, next\_q, dt)\;
        \popline \eIf{err <= TOL}{
            es $\gets$ euler\_segment(th0, th1, last\_p, next\_p)\;
            \tcp{Find location of any cusp in outline.}
            \tcp{If $0 \leq t < 1$, draw evolute at 't'.}
            cusp0 $\gets$ es\_curvature(es, 0) * offset + chord\_len\;
            cusp1 $\gets$ es\_curvature(es, 1) * offset + chord\_len\;
            t $\gets$ cusp0 * cusp1 >= 0 ? 1 : cusp0 / (cusp0 - cusp1)\;
            \If{cusp0 >= 0}{
                evolute\_finalize()\;
                es\_seg\_flatten\_offset(es, offset, vec2(0, t))\;
            }
            \If{cusp0 < 0 OR t < 1}{
                evolute\_start(es, offset)\;
                range $\gets$ cusp0 >= 0 ? vec2(t, 1) : vec2(0, t)\;
                es\_seg\_flatten\_evolute(es, range)\;
                es\_seg\_flatten\_offset\_reverse(es, offset, range)\;
            }
            \If{cusp0 < 0 AND t < 1}{
                evolute\_finalize()\;
                es\_seg\_flatten\_offset(es, offset, vec2(t, 1))\;
            }
            \tcp{Pop the stack}
            t0\_u $\gets$ t0\_u + 1\;
            shift $\gets$ countTrailingZeros(t0\_u)\;
            t0\_u $\gets$ t0\_u >> shift;
            dt $\gets$ dt * f32(1 << shift)\;
            last\_p $\gets$ next\_p\;
            last\_q $\gets$ next\_q\;
        }{
            \tcp{Error is above the tolerance.}
            \tcp{Push the stack; Continue subdivision}
            t0\_u $\gets$ t0\_u * 2\;
            dt $\gets$ dt * 0.5\;
        }
    }
}
\caption{Outline of the iterative cubic Bézier flattening routine invoked in our compute shader. A working WGSL implementation is provided in the supplemental materials.}
\label{alg:flatten-cubic}
\end{algorithm}
